Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-22 latex/main.tex
Updates from yan and shing

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{recombinant_evidence.pdf}
\caption{
Evidence for recombination in the ARG. All plots show ``robust'' sc2ts recombination nodes,
colored according to their classification by rebar
(recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations
when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labeled starting with `X' and
those identified by Jackson et al. \cite{Jackson2021} are prefixed with a `J', e.g.
Jackson Group B (JGB) and Jackson Singleton recombinant 1 (JS1).
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted
by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, Xx (the origin of XZ+XAC+XAD+XAE+XAP), XBB, and an undesignated recombination event under XBB, which we here refer to as  XBB.x; See Document S1.XXX for a description of these SnpIt-like\cite{xxx} plots, in which
each recombination node is a mix of variants matching the left parent (green), the right parent (blue), and de-novo mutations (gold).}
\label{fig:averted_mutations}
\end{figure}

@@

https://github.com/jeromekelleher/sc2ts-paper/issues/358
https://github.com/jeromekelleher/sc2ts-paper/issues/348

@@

\begin{figure}
\centering
% NOTE the size of the second "m" column doesn't matter, but we have to fix the
% size in order for vertical centering to work :sigh:
% The layout here isn't ideal, but it's incredibly fiddly to get center
% alignment veritically with cells to work, so this is good enough.
\begin{tabularx}{0.85\textwidth}{>{\centering\arraybackslash}m{1cm}>{\arraybackslash}m{1cm}}
Q1 & \includegraphics[height=\cprheight]{static/RE_node-QCpass-427863.png}\\
Q2 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-748991.png}\\
Q3 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-411345.png}\\
Q4 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-663484.png}\\
\end{tabularx}
\caption{
Illustrative copying patterns.
These copying patterns are examples
drawn from the quadrants labeled in Figure~\ref{fig:recombinant_qc}.
Each pattern indicates the genomic positions
where the allelic state in a recombinant
(middle row, with the label showing the node ID of the recombination event)
matches that of the suggested left or right parent (P0 and P1, respectively)
or where the allelic state differs from that of both the parents.
For reference, the allelic states of the Wuhan-Hu-1 genome
(which is the assumed root of the ARG)
are shown on top of the left parent.
The allelic states are colored as follows:
green for those matching the left parent;
blue for those matching the right parent;
yellow for those differing from both the parents; and
grey for those matching the reference.
Adjacent sites are underlined in red, and
near-adjacent sites (within 3 bases) in orange.
A continuous red+orange underline thus indicates
a region treated as a single locus.
For further context, mutations above a recombination event
are shown on the bottom.
}
\label{fig:copying_pattern_examples}
\end{figure}

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{recombinant_evidence.pdf}
\caption{
Evidence for recombination in the ARG. All plots show ``robust'' sc2ts recombination nodes,
colored according to their classification by rebar
(recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations
when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labeled starting with `X' and
those identified by Jackson et al. \cite{Jackson2021} are prefixed with a `J', e.g.
Jackson Group B (JGB) and Jackson Singleton recombinant 1 (JS1).
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted
by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, Xx (the origin of XZ+XAC+XAD+XAE+XAP), XBB, and an undesignated recombination event under XBB, which we here refer to as  XBB.x; See Fig~\ref{fig:copying_pattern_examples} for a description of these SnpIt-like~\cite{OToole2024} plots, in which
recombinants are shown  as a mosaic of variants inherited from the left vs right parents (green vs blue, de-novo mutations in gold).}
\label{fig:averted_mutations}
\end{figure}

@@

https://github.com/jeromekelleher/sc2ts-paper/issues/348
https://github.com/jeromekelleher/sc2ts-paper/issues/358

@@

\begin{figure}
\centering
% NOTE the size of the second "m" column doesn't matter, but we have to fix the
% size in order for vertical centering to work :sigh:
% The layout here isn't ideal, but it's incredibly fiddly to get center
% alignment veritically with cells to work, so this is good enough.
\begin{tabularx}{0.85\textwidth}{>{\centering\arraybackslash}m{1cm}>{\arraybackslash}m{1cm}}
Q1 & \includegraphics[height=\cprheight]{static/RE_node-QCpass-427863.png}\\
Q2 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-748991.png}\\
Q3 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-411345.png}\\
Q4 & \includegraphics[height=\cprheight]{static/RE_node-QCfail-663484.png}\\
\end{tabularx}
\caption{
Illustrative copying patterns
drawn from the quadrants labeled in Figure~\ref{fig:recombinant_qc}. Q1-Q4 correspond to copying patterns for recombination nodes 427863, 728991, 411345, and 663484 (none of which are associated with designated Pango recombinant lineages).
The SnpIt-like~\cite{OToole2024} schematic highlights the set of genomic positions
where the allelic state in a recombinant
(middle row, labelled by node ID)
matches that of the suggested left (P0, above, coloured green) or right parent (P1, below, coloured blue),
or where the recombination event requires a de-novo mutation (gold).
Genomic positions are shown above each column, with the allelic state of the Wuhan-Hu-1 reference genome (``ref'') immediately below;  parental states that are not inherited by the recombinant and correspond to the reference are shown with a grey background.
At the Adjacent genomic positions are underlined in red, and
near-adjacent sites (within 3 bases) in orange.
A continuous red+orange underline thus indicates
a region treated as a single locus.
For further context, mutations above a recombination event
are shown on the bottom.
}
\label{fig:copying_pattern_examples}
\end{figure}
